The quench action approach for integrable models: A Monte Carlo study

### 2015-11-02 draft/fs_qa_v4.tex
rev-latest

Understanding the out-of-equilibrium dynamics in {\it isolated} quantum 
many-body systems is one of the most intriguing research topics in 
contemporary physics, both theoretically~\cite{polkovnikov-2011} and 
experimentally~\cite{bloch-2008,greiner-2002,kinoshita-2006,
hofferberth-2007,trotzky-2012,gring-2012,cheneau-2012,
schneider-2012,kunhert-2013,langen-2013,meinert-2013,fukuhara-2013,
ronzheimer-2013,braun-2014,langen-2015}. The most popular out-of-equilibrium 
protocol is that of the {\it quantum quench} in which a system is initially 
prepared in an eigenstate $|\Psi_0\rangle$ of a many-body hamiltonian ${\mathcal H}_0$. 
Then a global parameter is suddenly changed, and the system is let to evolve 
unitarily under a new hamiltonian ${\mathcal H}$. There is now compelling 
evidence that at long times after the quench an equilibrium steady state 
arises (see for instance Ref.~\cite{kinoshita-2006}).

In the thermodynamic limit, due to dephasing, it is natural to expect that the 
equilibrium value of any local observable ${\mathcal O}$ is described by the 
so-called diagonal ensemble as 
%
\begin{equation}
\label{d-ensemble}
\langle{\mathcal O}\rangle_{DE}=\sum\limits_{\lambda}|\langle\Psi_0|\lambda
\rangle|^2\langle\lambda|{\mathcal O}|\lambda\rangle,
\end{equation}
%
where the sum is over the eigenstates $|\lambda\rangle$ of the post-quench 
hamiltonian ${\mathcal H}$. For generic, i.e., non-integrable, models 
the Eigenstate Thermalization Hypothesis~\cite{deutsch-1991,srednicki-1994} 
(ETH) implies that the diagonal ensemble~\eref{d-ensemble} becomes equivalent 
to the usual Gibbs (thermal) ensemble. 
%, implying 
%
%\begin{equation}
%\langle{\mathcal O}\rangle_{DE}=\langle{\mathcal O}\rangle_{Gibbs}\equiv
%\frac{1}{Z}\Tr({\mathcal O}\rho^{Gibbs}),\quad\textrm{with}\quad\rho^{Gibbs}
%=e^{-\beta{\mathcal H}},
%\end{equation}
%
%where $\beta$ is the inverse temperature of the system and $Z$ a normalization 
%factor. 
In integrable models, however, the presence of an extensive number of local or 
quasi-local integrals of motion strongly affects the dynamics, preventing the onset of 
thermal behavior. It has been suggested in Ref.~\cite{rigol-2007,rigol-2008} that   
in this situation the post-quench steady-state can be described by the so-called 
Generalized Gibbs Ensemble ($GGE$) as 
%
\begin{equation}
\label{gge}
\langle{\mathcal O}\rangle_{GGE}=\frac{1}{Z}\Tr({\mathcal O}\rho^{GGE}), 
\quad\textrm{with}\quad\rho^{GGE}\equiv\frac{1}{Z}\exp\Big(-\sum_j\beta_jQ_j\Big). 
\end{equation}
%
Here $Q_j$ are mutually commuting conserved quantities, i.e., $[{\mathcal H},Q_j]=0\,
\forall j$ and $[Q_j,Q_k]=0\,\forall j,k$, whereas $\beta_j$ are Lagrange multipliers 
to be fixed by imposing that  $\langle Q_j\rangle_{GGE}=\langle\Psi_0|Q_j|\Psi_0\rangle$. 
The validity of the GGE has been confirmed in non-interacting theories~\cite{calabrese-2011,
calabrese-2012,fagotti-2013,kcc14,kcc14a}, whereas in interacting ones the scenario is 
still not settled. A recent work~\cite{ilievski-2015a} suggests that the $GGE$ 
description of the steady state  is complete provided that the so-called quasi-local 
charges~\cite{prosen-2014,pereira-2014,ilievski-2015} are included in~\eref{gge}.

Valuable insights into this issue have been provided by the so-called quench-action 
approach~\cite{caux-2013}. For Bethe ansatz integrable models this allows to construct 
the diagonal ensemble average~\eref{d-ensemble} in the thermodynamic limit, provided 
that the overlaps $\langle\lambda|\Psi_0\rangle$ are known. The physical idea is that 
for large system sizes it is possible to approximate the sum over the eigenstates 
in~\eref{d-ensemble} using a saddle point argument. Specifically, one starts with 
rewriting~\eref{d-ensemble} as 
%
\begin{equation}
\label{de1}
\langle{\mathcal O}\rangle_{DE}= \sum\limits_{\lambda}\rho^{DE}\langle\lambda|
{\mathcal O}|\lambda\rangle,\quad\textrm{with}\quad\rho^{DE}=\exp(-2\Re{\mathcal 
E}(\lambda)), 
\end{equation}
%
where ${\mathcal E}(\lambda)\equiv-\log\langle\lambda|\Psi_0\rangle$, and $\Re$ 
denoting the real part. For typical initial states $|\Psi_0\rangle$
one has ${\mathcal E}\propto L$, with $L$ the system size, reflecting  a typical 
vanishing of the overlaps as $\langle\Psi_0|\lambda\rangle\propto e^{-c L}$. 
Similar to the standard Thermodynamic Bethe Ansatz (TBA)~\cite{taka-book}, the 
extensivity of 
${\mathcal E}$ suggests that in the thermodynamic limit~\eref{de1} is dominated 
by a saddle point. Surprisingly, for some Bethe ansatz solvable models and some 
simple initial states standard TBA techniques, 
complemented with analytical knowledge of the overlaps $\langle\Psi_0|\lambda
\rangle$, allow to determine {\it analytically} this saddle point, and for some  
local observables, the diagonal ensemble expectation values~\eref{de1}, in the 
thermodynamic limit. This has been done succesfully for instance in the Lieb-Liniger 
model for the quench from the Bose-Einstein condensate~\cite{de-nardis-2014} 
(BEC) and for the $XXZ$ spin chain for the quench from the N\'eel state~\cite{wouters-2014,
brockmann-2014}. Interestingly, recently it has been suggested that the quench action 
approach could allow to reconstruct the full relaxation dynamics to the steady 
state~\cite{de-nardis-2015a,de-nardis-2015}.

%Formally,~\eref{de1} is equivalent to the standard Gibbs 
%ensemble after replacing ${\mathcal H}\to{\mathcal E}$. A crucial difference is 
%that while typical hamiltonians are local, ensuring the extensivity of the energy, 
%this is not guaranteed for ${\mathcal E}$.

Here we provide a Monte Carlo implementation of the quench action approach in the 
finite-size spin-$1/2$ Heisenberg chain ($XXX$ chain). combining Monte Carlo methods and Bethe ansatz techniques.

We focus on the situation in which the pre-quench initial state is the Neel 
state or the Majumdar-Ghosh state.

We investigate the importance of the zero-momentum strings in the Quench Action.

Without zero-momentum strings the overlap saturation rules are not valid, 
i.e., in finite size systems the vast majority of the eigenstates contain 
zero momentum strings.

The details on the eigenstates counting depend on the pre-quench initial state.

However, we show that one can restrict to the set of non-zero momentum strings. 
The fact that one neglects zero-momentum strings gives rise only to scaling 
corrections.

We also investigate the validity of the Bethe-Takahashi approximation for the 
calculation of the overlap.

@@

The N\'eel state is defined as 
%
\begin{equation}
\label{neel}
|N\rangle\equiv\frac{1}{\sqrt{2}}\big(|N_1\rangle
+|N_2\rangle\big), 
\end{equation}
%
with $|N_1\rangle\equiv |\uparrow\downarrow\rangle^{\otimes L/2}$, and $|N_2\rangle\equiv 
|\downarrow\uparrow\rangle^{\otimes L/2}$. Note that $|N_1\rangle=\hat T|N_2\rangle$, 
with $\hat T$ the one-site shift operator.

The Majumdar-Ghosh state $|MG\rangle$ is defined as 
%
\begin{equation}
\label{mg}
|MG\rangle\equiv \Big(\frac{|\uparrow\downarrow\rangle-|\downarrow\uparrow\rangle}
{\sqrt{2}}\Big)^{\otimes L/2}. 
\end{equation}
%

@@

Here we review some Bethe ansatz results for the spin-$\frac{1}{2}$ Heisenberg 
($XXX$) chain. Specifically, in section~\ref{sec:1.1} we introduce the model. 
Its eigenstates (Bethe states) and the Bethe equations are discussed in 
section~\ref{sec:1.2}. Section~\ref{sec:1.3} focuses on the string hypothesis 
and the so-called Bethe-Gaudin-Takahashi (BGT) equations. The form of the 
BGT equations in the thermodynamic limit is discussed in section~\ref{sec:1.4}. 
Finally, in section~\ref{sec:1.5} we provide the exact formulas for the 
local conserved charges of the model.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The spin-$\frac{1}{2}$ Heisenberg chain}
\label{sec:1.1}

The spin-$\frac{1}{2}$ isotropic Heisenberg chain ($XXX$ chain) with $L$ sites 
is defined by the Hamiltonian 
%
\begin{equation}
\label{xxx-ham}
{\mathcal H}\equiv J\sum\limits_{i=1}^L\left[\frac{1}{2}(S_i^+S^-_{i+1} 
+S_i^{-}S_{i+1}^+)+S_i^zS_{i+1}^z-\frac{1}{4}\right],  
\end{equation}
%
where $S^{\pm}_i\equiv (\sigma_i^x\pm i\sigma_i^y)/2$ are spin operators acting on the 
site $i$, $S_i^z\equiv\sigma_i^z/2$, and $\sigma^{x,y,z}_i$ the Pauli matrices. Here we 
fix $J=1$ and use periodic boundary conditions, identifying sites $L+1$ and $1$. The 
total magnetization $S_{T}^z\equiv\sum_iS_i^z=L/2-M$, with $M$ number of down spins 
(particles), commutes with~\eref{xxx-ham}, and it is used to label its eigenstates.

@@

It is interesting to consider the situation with $M/L\to\infty$ (low-density limit). 
We first consider the case with all the particles forming a single zero-momentum 
string. Due to parity invariance only strings of even length are allowed. 
We assume the string hypothesis~\eref{str-hyp}, restricting ourselves to zero-momentum 
strings, i.e., 
%
\begin{equation}
\lambda^{j}_{n}=-i(n-1-2j)+i\delta_{n}^j,
\end{equation}
%
with $n$ the length of the string and $j$ labelling the different string components. 
Due to parity invariance we restrict ourselves to half strings, i.e., $j=1,\dots,n/2$.
We verified by explicit computation that for a single $m$ string one has (at leading order 
in the string deviations) 
%
\begin{equation}
\det(G^+)=-\frac{2^{n/2}}{\delta_n^{n/2}}L\prod_{j>k=1}^{n/2-1}(\delta_n^k-\delta_n^j)^{-1},
\end{equation}
%
and 
%
\begin{equation}
\det(G^-)=-\frac{2^{n/2}}{\delta_n^{n/2}}(L-1)\prod_{j>k=1}^{n/2-1}(\delta_n^k-\delta_n^j)^{-1}.
\end{equation}
%
This implies that 
%
\begin{equation}
\frac{\det_n(G^+)}{\det_n(G^-)}\to\frac{L}{L-1},\quad\forall n.
\end{equation}
%
However, it is straightforward to check that  
%
\begin{equation}
\prod_{j=1}^{n/2}\frac{\sqrt{\lambda_j^2+1}}{4\lambda_j}\propto\sqrt{\delta_n^{n/2}}.
\end{equation}
%
{\bf The prefactor does not seem to be a simple function of $n$}.

@@

A problem arises if one considers the Majumdar-Ghosh state. The overlap is obtained 
rescaling the N\'eel overlap (see below) with the prefactor
%
\begin{equation}
\prod\limits_{j=1}^n\frac{1}{2}
\Big(1-\frac{\lambda_j-i}{\lambda_j+i}\Big). 
\end{equation}
%
Note that the product above is over all the string components. 
For a single string one gets 
%
\begin{equation}
\prod\limits_{j=1}^n\frac{1}{2}
\Big(1-\frac{\lambda_j-i}{\lambda_j+i}\Big)\propto\frac{1}{\delta_n^{n/2}}.
\end{equation}
%
Putting everything together one would get a diverging MG overlap which is clearly wrong.

>>>>

Understanding the out-of-equilibrium dynamics in {\it isolated} quantum 
many-body systems is one of the most intriguing research topics in 
contemporary physics, both experimentally~\cite{bloch-2008,greiner-2002,
kinoshita-2006,hofferberth-2007,trotzky-2012,gring-2012,cheneau-2012,
schneider-2012,kunhert-2013,langen-2013,meinert-2013,fukuhara-2013,
ronzheimer-2013,braun-2014,langen-2015} and theoretically~\cite{polkovnikov-2011}. 
The most investigated protocol is that of the {\it quantum quench}, in 
which a system is initially prepared in an eigenstate $|\Psi_0\rangle$ of 
a many-body hamiltonian ${\mathcal H}_0$. Then a global parameter is 
suddenly changed, and the system is let to evolve unitarily under a new 
hamiltonian ${\mathcal H}$. There is now compelling evidence that at long 
times after the quench an equilibrium steady state arises (see for instance 
Ref.~\cite{kinoshita-2006}), although its nature is not fully understood yet. 
In the thermodynamic limit, due to dephasing, it is natural to expect that 
the equilibrium value of any local observable ${\mathcal O}$ is described 
by the so-called diagonal ensemble as 
%
\begin{equation}
\label{d-ensemble}
\langle{\mathcal O}\rangle_{DE}=\sum\limits_{\lambda}|\langle\Psi_0|\lambda
\rangle|^2\langle\lambda|{\mathcal O}|\lambda\rangle,
\end{equation}
%
where the sum is over the eigenstates $|\lambda\rangle$ of the post-quench 
hamiltonian ${\mathcal H}$. Moreover, for generic, i.e., non-integrable, models 
the Eigenstate Thermalization Hypothesis~\cite{deutsch-1991,srednicki-1994} 
(ETH) implies that the diagonal ensemble~\eref{d-ensemble} becomes equivalent 
to the usual Gibbs (thermal) ensemble.

In integrable models, however, the presence of an extensive number of local or 
quasi-local integrals of motion strongly affects the dynamics, preventing the 
onset of thermal behavior. It has been suggested in Ref.~\cite{rigol-2007,
rigol-2008} that in this situation the post-quench steady-state can be described 
by the so-called Generalized Gibbs Ensemble ($GGE$) as 
%
\begin{equation}
\label{gge}
\langle{\mathcal O}\rangle_{GGE}=\frac{1}{Z}\Tr({\mathcal O}\rho^{GGE}), 
\quad\textrm{with}\quad\rho^{GGE}\equiv\frac{1}{Z}\exp\Big(-\sum_j\beta_jQ_j\Big). 
\end{equation}
%
Here $Q_j$ are mutually commuting conserved quantities, i.e., $[{\mathcal H},Q_j]=0\,
\forall j$ and $[Q_j,Q_k]=0\,\forall j,k$, whereas $\beta_j$ are Lagrange multipliers 
to be fixed by imposing that  $\langle Q_j\rangle_{GGE}=\langle\Psi_0|Q_j|\Psi_0\rangle$. 
Note that for the diagonal ensemble~\eref{d-ensemble}  one has trivially that $\langle
{\mathcal O}\rangle_{DE}=\langle\Psi_0|{\mathcal O}|\Psi_0\rangle$. We anticipate that 
this provides  useful sum rules that will be the main object of interest in the paper. 
The validity of the GGE has been confirmed in non-interacting theories~\cite{calabrese-2011,
calabrese-2012,fagotti-2013,kcc14,kcc14a}, whereas in interacting ones the scenario is 
still not settled. Note that a recent work~\cite{ilievski-2015a} suggests that the $GGE$ 
description of the steady state  is complete provided that the so-called quasi-local 
charges~\cite{prosen-2014,pereira-2014,ilievski-2015} are included in~\eref{gge}.

Valuable insights into this issue have been provided by the so-called quench-action 
approach~\cite{caux-2013}. For Bethe ansatz integrable models this method allows one 
to construct the diagonal ensemble~\eref{d-ensemble} directly in the thermodynamic 
limit, provided that the overlaps $\langle\lambda|\Psi_0\rangle$ are known. The 
physical idea is that for large system sizes it is possible to approximate the sum 
over the eigenstates in~\eref{d-ensemble} using a saddle point argument. Specifically, 
one starts with rewriting~\eref{d-ensemble} as 
%
\begin{equation}
\label{de1}
\langle{\mathcal O}\rangle_{DE}= \sum\limits_{\lambda}\rho^{DE}\langle\lambda|
{\mathcal O}|\lambda\rangle,\quad\textrm{with}\quad\rho^{DE}=\exp(-2\Re{\mathcal 
E}(\lambda)), 
\end{equation}
%
where ${\mathcal E}(\lambda)\equiv-\log\langle\lambda|\Psi_0\rangle$, and $\Re$ 
denotes the real part. For typical initial states $|\Psi_0\rangle$
one has ${\mathcal E}\propto L$, with $L$ the system size. This reflects the 
vanishing of the overlaps as $\langle\Psi_0|\lambda\rangle\propto e^{-c L}$.  
As in the standard Thermodynamic Bethe Ansatz (TBA)~\cite{taka-book}, the 
extensivity of ${\mathcal E}$ suggests that in the thermodynamic limit the 
sum in~\eref{de1} is dominated by a saddle point. Remarkably, for some Bethe 
ansatz solvable models and for simple-enough initial states, it is possible to 
determine this saddle point {\it analytically}. This has been done succesfully, 
for instance, for the quench from the Bose-Einstein condensate~\cite{de-nardis-2014} 
(BEC) in the Lieb-Liniger model and for the quench from the N\'eel 
state~\cite{wouters-2014,brockmann-2014} in the $XXZ$ spin chain. Interestingly, 
it has been suggested recently that the quench action approach could allow to 
reconstruct the full relaxation dynamics to the steady state~\cite{de-nardis-2015a,
de-nardis-2015}.

\paragraph*{Outline of the results.} 
In this paper, by combining exact Bethe ansatz and Monte Carlo techniques, we investigate 
the diagonal ensemble and the quench action approach in {\it finite} size integrable models. 
We focus on the spin-$1/2$ isotropic Heisenberg ($XXX$) chain with $L$ sites, which is 
defined by the Hamiltonian 
%
\begin{equation}
\label{xxx-ham}
{\mathcal H}\equiv J\sum\limits_{i=1}^L\left[\frac{1}{2}(S_i^+S^-_{i+1} 
+S_i^{-}S_{i+1}^+)+S_i^zS_{i+1}^z-\frac{1}{4}\right].
\end{equation}
%
Here $S^{\pm}_i\equiv (\sigma_i^x\pm i\sigma_i^y)/2$ are spin operators acting on the 
site $i$, $S_i^z\equiv\sigma_i^z/2$, and $\sigma^{x,y,z}_i$ the Pauli matrices. We 
fix $J=1$ in~\eref{xxx-ham} and use periodic boundary conditions, identifying sites 
$L+1$ and $1$. The total magnetization $S_{T}^z\equiv\sum_iS_i^z=L/2-M$, with $M$ number 
of down spins (particles), commutes with~\eref{xxx-ham}, and it is used to label its 
eigenstates. The $XXX$ chain is Bethe ansatz solvable~\cite{bethe-1931} and its eigenstates 
are in one-to-one correspondence with the solutions of the so-called Bethe equations 
(see~\ref{sec:1.2}). We restrict ourselves to quenches from the zero-momentum N\'eel 
state $|N\rangle$ and the Majumdar-Ghosh state $|MG\rangle$, which are defined as  
%
\begin{equation}
\label{psi0}
|N\rangle\equiv\Big(\frac{|N_1\rangle
+|N_2\rangle}{\sqrt{2}}\Big), \qquad 
|MG\rangle\equiv \Big(\frac{|\uparrow\downarrow\rangle-|\downarrow\uparrow\rangle}
{\sqrt{2}}\Big)^{\otimes L/2}. 
\end{equation}
%
Here $|N_1\rangle\equiv |\uparrow\downarrow\rangle^{\otimes L/2}$, and $|N_2\rangle
\equiv |\downarrow\uparrow\rangle^{\otimes L/2}$, with $\otimes$ denoting the tensor 
product. Note that both $|N\rangle$ and $|MG\rangle$ are invariant under one site 
translations. Our study relies on the analytical knowledge of the overlaps between 
the N\'eel and Majumdar-Ghosh state and the eigenstates of~\eref{xxx-ham}~\cite{
brockmann-2014,brockmann-2014a,brockmann-2014b,brockmann-2014c,pozsgay-2014a,
piroli-2014}.

We first present a detailed overview of the distribution of the overlaps between 
the initial states and the eigenstates of the model (overlap distribution function). 
Precisely, we provide numerical results for all the overlaps for finite chains up 
to $L\le 38$. Our results are obtained exploiting the analytical formulas 
presented in Ref.~\cite{brockmann-2014} and~\cite{pozsgay-2014a}. Crucially, 
we restrict ourselves to a truncated Hilbert space, considering only eigenstates 
of~\eref{xxx-ham} that do not contain zero-momentum strings. Physically, 
zero-momentum strings correspond to eigenstates amplitudes containing multi-particle  
bound states which are fully delocalized in the chain. From the Bethe ansatz perspective, 
in the thermodynamic limit the presence of zero-momentum strings leads to fictitious 
singularities in the overlap formulas. Dealing with these singularities is 
a formidable task that requires detailed knowledge of the solutions of the 
Bethe equations, and it can be done only in very simple cases, e.g., for small 
chains. On the other hand, it has been argued that zero-momentum strings are irrelevant 
in the thermodynamic limtit~\cite{brockmann-2014a}.

For a finite chain here we find that for both the N\'eel and the Majumdar-Ghosh 
cases, the fraction of eigenstates that do not contain zero-momentum strings 
is vanishing in the thermodynamic limit, meaning that for finite large-enough 
systems the vast majority of the eigenstates contain zero momentum strings. Specifically, 
the total number of eigenstates without zero-momentum strings is given in terms of the 
chain size $L$ by simple combinatorial formulas that we provide. We investigate the 
effect of the Hilbert space truncation on the diagonal ensemble focusing on the  
sum rules $\langle Q_j\rangle_{DE}=\langle\Psi_0|Q_j|\Psi_0\rangle$ for the conserved 
quantities $Q_j$. We numerically demonstrate that truncating the diagonal ensemble to 
eigenstates with no zero-momentum strings leads to striking violations of the sum 
rules. Precisely, we numerically observe that the truncated diagonal ensemble 
average $\langle Q_j\rangle$ vanish in the thermodynamic limit reflecting the 
vanishing behavior of the fraction of finite-momentum strings eigenstates. 
This allows us to conclude that for finite chaings eigenstates corresponding 
to zero-momentum strings cannot be trivially neglected when considering 
diagonal ensemble averages.

However, we provide numerical evidence that the correct diagonal ensemble 
sum rules can be recovered in the thermodynamic limit by appropriately 
reweighting the contribution of the finite-momentum strings eigenstates.  
We focus only on the situation with the N\'eel state (cf.~\eref{psi0}) as 
initial state. The key idea is to sample the eigenstates with no zero-momentum 
strings with the diagonal ensemble probability measure $\rho^{DE}$ (cf.~\eref{de1}). 
To this purpose we develop a Monte Carlo scheme, which is a generalization of 
the approach presented in Ref.~\cite{alba-2015} to simulate the Generalize 
Gibbs Ensemble. The method is based on the knowledge of the N\'eel overlaps 
$\langle\lambda|N\rangle$ and on the knowledge of the Hilbert space structure 
of the $XXX$ chain in the Bethe ansatz formalism. The approach allowed us to 
simulate effectively chain with $L\lesssim 60$, although larger systems 
sizes can in principle be reached.

Strikingly, although for small chains violations of the conserved quantities 
sum rules are present, these violations vanish in the thermodynamic limit 
and the sum rules are restored. This implies that the only effect of the 
Hilbert space truncation is to introduce scaling corrections. In the quench 
action language, this means that the eigenstates corresponding to no 
zero-momentum strings contain enough physical information about saddle point. 
This is numerically confirmed by extracting the so-called saddle point 
root distributions, which in the Bethe ansatz language fully characterize 
the diagonal ensemble averages in the thermodynamic limit. In the numerical  
 approach these are obtained from the histograms of the Bethe ansatz 
solutions sampled during the Monte Carlo. A part from finite size scaling 
corrections, we observe striking agreement with analytical results, at least 
for the first few root distributions.

@@

%The N\'eel state is defined as 
%
%\begin{equation}
%\label{neel}
%|N\rangle\equiv\frac{1}{\sqrt{2}}\big(|N_1\rangle
%+|N_2\rangle\big), 
%\end{equation}
%

%The Majumdar-Ghosh state $|MG\rangle$ is defined as 
%
%\begin{equation}
%\label{mg}
%|MG\rangle\equiv \Big(\frac{|\uparrow\downarrow\rangle-|\downarrow\uparrow\rangle}
%{\sqrt{2}}\Big)^{\otimes L/2}. 
%\end{equation}
%

@@

Here we review some Bethe ansatz results for the spin-$\frac{1}{2}$ Heisenberg 
($XXX$) chain. Specifically, in section~\ref{sec:1.2} discuss the structure of  
its eigenstates (Bethe states) and the associated Bethe equations. 
Section~\ref{sec:1.3} focuses on the string hypothesis 
and the so-called Bethe-Gaudin-Takahashi (BGT) equations. The form of the 
BGT equations in the thermodynamic limit is discussed in section~\ref{sec:1.4}. 
Finally, in section~\ref{sec:1.5} we provide the exact formulas for the 
local conserved charges of the model.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{The spin-$\frac{1}{2}$ Heisenberg chain}
%\label{sec:1.1}

@@

It is interesting to consider the situation with $M/L\to\infty$ (low-density limit). 
We first consider the case with all the particles forming a single zero-momentum 
string. Due to parity invariance only strings of even length are allowed. 
We assume the string hypothesis~\eref{str-hyp}, restricting ourselves to zero-momentum 
strings, i.e., 
%
\begin{equation}
\lambda^{j}_{n}=-i(n-1-2j)+i\delta_{n}^j,
\end{equation}
%
with $n$ the length of the string and $j$ labelling the different string components. 
Due to parity invariance we restrict ourselves to half strings, i.e., $j=1,\dots,n/2$.
We verified by explicit computation that for a single $m$ string one has (at leading order 
in the string deviations) 
%
\begin{equation}
\det(G^+)=-\frac{2^{n/2}}{\delta_n^{n/2}}L\prod_{j>k=1}^{n/2-1}(\delta_n^k-\delta_n^j)^{-1},
\end{equation}
%
and 
%
\begin{equation}
\det(G^-)=-\frac{2^{n/2}}{\delta_n^{n/2}}(L-1)\prod_{j>k=1}^{n/2-1}(\delta_n^k-\delta_n^j)^{-1}.
\end{equation}
%
This implies that 
%
\begin{equation}
\frac{\det_n(G^+)}{\det_n(G^-)}\to\frac{L}{L-1},\quad\forall n.
\end{equation}
%
However, it is straightforward to check that  
%
\begin{equation}
\label{wrong}
\prod_{j=1}^{n/2}\frac{\sqrt{\lambda_j^2+1}}{4\lambda_j}\propto\sqrt{\delta_n^{n/2}}.
\end{equation}
%
{\bf The prefactor does not seem to be a simple function of $n$}.

@@

A problem arises if one considers the Majumdar-Ghosh state. The overlap is obtained 
rescaling the N\'eel overlap (see below) with the prefactor
%
\begin{equation}
\prod\limits_{j=1}^n\frac{1}{2}
\Big(1-\frac{\lambda_j-i}{\lambda_j+i}\Big). 
\end{equation}
%
Note that the product above is over all the string components. 
For a single string one gets 
%
\begin{equation}
\label{wrong1}
\prod\limits_{j=1}^n\frac{1}{2}
\Big(1-\frac{\lambda_j-i}{\lambda_j+i}\Big)\propto\frac{1}{\delta_n^{n/2}}.
\end{equation}
%
Putting everything together (\eref{wrong} and~\eref{wrong1}) one would get a 
diverging MG overlap, which is clearly impossible.
